---
title: "Linear Regression"
author: "Coreys Notes"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 3
---

```{r libraries, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(caret)
options(scipen = 999)
```

```{r importing}
train <- read.csv("train.csv",stringsAsFactors = FALSE)
train <- train %>% 
  select_if(is.numeric)
mymod <- lm(SalePrice~.,data=train,na.action = "na.omit")
```


```{r conf and pred intervals}
confint(mymod,level = .95)
#predict(mymod,data.frame(x1=.,x2=.,x3=...),interval="confidence"or"prediction")

```

# Interaction Terms
```{r interaction terms}
# add one interaction term

oneinteraction <- lm(SalePrice~MSSubClass*LotFrontage+.,data=train)
oneinteraction

lminteractions <- lm(SalePrice~.^2,data=train)
#summary(lminteractions)

```

# Polynomial Terms and Log

```{r polynomial terms}

polymod <- lm(SalePrice~poly(LotArea,2),data=train)
summary(polymod)

# log of predictor
logmod <- lm(SalePrice~log(LotArea),data=train)
summary(logmod)


```

# Partial F test

tests if model with added predictors is better. Statistically significant means better.

```{r partial f test}
mymod1 <- lm(SalePrice~LotArea,data=train)
mymod2 <- lm(SalePrice~poly(LotArea,2),data=train)
anova(mymod1,mymod2)
```

# Assumptions of Linear Regression

## Outliers
```{r outliers}
library(MASS,include.only = "studres")
studentized <- lm(SalePrice~LotArea,data=train)
studResiduals <- studres(studentized)
head(studResiduals)
studDF <- data.frame(predictor=train$LotArea,studResid=studResiduals)
ggplot(studDF,aes(x=predictor,y=studResid))+
  geom_point()

train <- cbind(train,stud=studDF$studResid)
names(train)
train %>% 
  select(stud,LotArea,SalePrice) %>% 
  arrange(stud) %>% 
  head()
```

## Cooks distance

```{r cooks distance}
plot(cooks.distance(lm(SalePrice ~ LotArea + LotFrontage, data=train)))

as.data.frame(cooks.distance(lm(SalePrice~.,data=train))) %>% 
  arrange(desc(.)) %>% 
  head()

```

## Multicollinearity

Variance inflation factors, above 5 is cause for concern, 10 is high. 
```{r variance inflation factors}
train <- read.csv("train.csv",stringsAsFactors = FALSE)
train <- train %>% select_if(is.numeric)
# wont run if high correlations
train <- train[,-findCorrelation(cor(train,use="complete.obs"),cutoff = .8)]

mymod <- lm(SalePrice~.,data=train,na.action="na.omit")
library(car)
as.data.frame(vif(mymod)) %>% 
  arrange(desc(.)) %>% 
  head()

```


## Correlations

Highly correlated predictors
```{r corr plot}
library(corrplot)
corrplot(cor(train,use="complete.obs"))

findCorrelation(cor(train,use="complete.obs"),cutoff = .8,verbose=TRUE)

train[,findCorrelation(cor(train,use="complete.obs"),cutoff = .8,verbose=TRUE)]

cor(train[,27],train[,28],use="complete.obs")
```


Finding all correlations between the independent variables and the dependent variable. 
```{r find correlations}
SalePriceCorrelations <- as.data.frame(sapply(train,function(x) cor(x,train$SalePrice,use="complete.obs")))

names(SalePriceCorrelations) <- c("correlation")

SalePriceCorrelations %>% 
  filter(abs(correlation)>.5) %>% 
  arrange(desc(correlation))

```

## High Leverage

```{r high leverage points}
# hatvalues are leverage values, or unusual predictor values

plot(hatvalues(mymod))
hatvalues(mymod) %>% 
  as.data.frame(.) %>% 
  arrange(desc(.)) %>% 
  head()
```

## Heteroscedasticity
```{r heteroscedasticity}
# if less than .05 then heteroscedasticity exists
library(lmtest)
bptest(mymod)
```
# Feature Selection Methods

## Forward Selection, backward, stepwise

```{r forward selection}
train <- train[complete.cases(train),]
#leapBackward, leapForward, leapSeq(stepwise)
library(caret)

ctrl <- trainControl(method="cv",number=10)
set.seed(147)
stepmodel <- train(SalePrice~.,data=train,
                   method="leapForward",
                   tuneGrid=expand.grid(nvmax=1:30),
                   trControl=ctrl)
stepmodel$bestTune
stepmodel
stepmodel$results
stepmodel$finalModel
summary(stepmodel$finalModel)
# coefficients of final model
coef(stepmodel$finalModel,20)

```