---
title: "PCA"
author: "Coreys Notes"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 3
---
```{r libraries, message=FALSE, warning=FALSE}

library("FactoMineR")
library("factoextra")
```

```{r}
data("decathlon2")
mydf <- decathlon2
head(mydf)
mydf <- mydf[1:23, 1:10]
myPCA <- PCA(mydf,scale.unit=TRUE,graph=FALSE)
myPCA

```

```{r eigen values}
eigValues <- get_eigenvalue(myPCA)
eigValues
```

# Scree Plot

One method for choosing the number of PC's to only keep eigenvalues above one. Another possible option is to keep enough PC's to account for at least 70% of the data. 

```{r scree plot}
fviz_eig(myPCA,addlabels = TRUE, ylim=c(0,50))
```

# Variable Correlation Plot

Positively correlated variables are grouped together.

Negatively correlated variables are positioned on opposite sides of the plot origin (opposed quadrants).

The distance between variables and the origin measures the quality of the variables on the factor map. Variables that are away from the origin are well represented on the factor map. 



```{r variable correlation plot}
var <- get_pca_var(myPCA)

head(var$coord,5)

fviz_pca_var(myPCA,col.var = "black")
```

From the graph shotput and x100m are negative correlated with each other. x100m and x110m hurdle are correlated with each othter and shotput and discus are correlated with each other. 


# PCR

```{r}
mydf2 <- data.frame(y=c(1,4,4,5,3,5,6,7,10,12),x1=c(4,2,3,5,6,4,6,7,7,5),x2=c(2,3,4,3,4,6,10,12,15,12))

library (pls)
set.seed(2)
pcr.fit <- pcr(y~.,data=mydf2,scale=TRUE, validation ="CV") # 10 fold cv

summary(pcr.fit)

# cross validation MSE plot
validationplot(pcr.fit ,val.type="MSEP")

# test error
pcr.pred <- predict(pcr.fit,mydf2[6:10,],ncomp =2)
mean((pcr.pred -mydf2[6:10,"y"])^2)

# perform on full data set for final model
#pcr.fit <- pcr(y∼x,scale=TRUE ,ncomp=7)
#summary (pcr.fit)
```


```{r}
states <- row.names(USArrests)
head(USArrests)
sapply(USArrests,function(x) var(x))
```


